Loan Data Analysis by Andrew Voorhees
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
#important that plyr is loaded before dplyr
library(plyr)
library(dplyr)
library(scales)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
data = read.csv('prosperLoanData.csv', na.strings=c("", "NA"))

```


This report explores a dataset containing over 100,000 loans from the lending
company Prosper. 

# Univariate Plots Section


```{r echo=FALSE, Univariate_Plots}
dim(data)

```

Although there are 81 fields, I want to concern myself with only a limited 
number of these for my analysis. The fields I chose to focus on were:

1. CreditGrade
2. ListingCreationDate
3. Term 
4. LoanStatus
5. BorrowerAPR
6. Prosper Rating
7. ProsperScore
8. ListingCategory
9. BorrowerState
10. CreditRangeLower / CreditRangeUpper
11. StatedMonthlyIncome
12. LoanOriginalAmount
13. Investors


```{r}
keep = c('CreditGrade', 
         'ListingCreationDate', 
         'Term',
         'BorrowerAPR',
         'ProsperRating..numeric.', 
         'ProsperScore',
         'ListingCategory..numeric.',
         'CreditScoreRangeLower',
         'CreditScoreRangeUpper',
         'StatedMonthlyIncome',
         'LoanOriginalAmount',
         'Investors')
data = data[keep]

str(data)
```

Next I want to do some basic data wrangling to organize my data. 

```{r}
# Want Credit Grade to be an ordered Factor
data$CreditGrade = factor(data$CreditGrade, 
                          ordered=T, 
                          levels=c("AA", "A", "B", "C",
                                   "D",  "E",  "HR", "NC"))
# Creation date
data$ListingCreationDate = as.Date(data$ListingCreationDate)
data$CreationMonth = as.Date(cut(data$ListingCreationDate, breaks='month'))

p_levels = unique(sort(format(as.Date(data$ListingCreationDate), '%m')))
data$NumericMonth = factor(format(as.Date(data$ListingCreationDate), 
                                          '%m'), 
                                   ordered=TRUE, 
                                   levels=p_levels)
# Term is probably more accurately reflected as an ordered factor as well
data$Term = factor(data$Term, ordered=TRUE, levels=c(12, 36, 60))
#Renaming ProsperRating and setting as a factor
names(data)[names(data) == 'ProsperRating..numeric.'] = 'ProsperRating'
data$ProsperRating = factor(data$ProsperRating, ordered=TRUE, levels=seq(1, 7, 1))
# Ordered factor for Prosper Score
data$ProsperScore = factor(data$ProsperScore, ordered=TRUE, levels=seq(1, 11, 1))
# Rename listing category and map to actual concepts
names(data)[names(data) == 'ListingCategory..numeric.'] = 'ListingCategory'
p_from = seq(0, 20, 1)
# this is frustrating, but I'm not sure there's a clearer way
p_to = c('Not Available', 
         'Debt Consolidation',
         'Home Improvement', 
         'Business',
         'Personal Loan', 
         'Student Use',
         'Auto',
         'Other',
         'Baby&Adoption',
         'Boat', 
         'Cosmetic Procedure', 
         'Engagement Ring', 
         'Green Loans',
         'Household Expenses',
         'Large Purchases',
         'Medical/Dental', 
         'Motorcycle',
         'RV', 
         'Taxes',
         'Vacation',
         'Wedding Loans')

data$ListingCategory = factor(mapvalues(data$ListingCategory, 
                                        from = p_from, 
                                        to = p_to), 
                              levels=p_to)
data$CreditScore = (data$CreditScoreRangeUpper + data$CreditScoreRangeLower) / 2
data$CreditScoreRangeUpper = NULL 
data$CreditScoreRangeLower = NULL 
```

Great, so what does our data look like now?

```{r}
str(data)
```


```{r}
summary(data)
```

# Univariate Analysis Plots

```{r}
ggplot(aes(x=Investors), data=data) + 
  geom_histogram(boundary=0)  
```

Intuitively, having a number of investors already paired with a loan might 
encourage further investment. This indicates that the number of investors might 
follow a multiplicative process. I'll take a log transformation to better grasp 
the distribution. 

```{r}
ggplot(aes(x=Investors), data=data) + 
  geom_histogram(boundary=0) + 
  scale_x_log10()

```

Interesting, so if you throw out the common case where there are zero
investors, the distribution appears roughly normal or skewed normal. Definitely 
want to look at that peak on the far left though


```{r}
sort(table(data$Investors), decreasing=TRUE)[1:5]
```

The peak at the beginning is one of the issues with taking logs of counts. 
When the # of investors is 1, the log of that value is 0 which screws 
with my distribution quite a bit. 


Next I want to look into what credit grade is assigned to most of the listings. 


```{r}
ggplot(aes(x=CreditGrade), data=subset(data, !is.na(CreditGrade))) + 
  geom_bar()
```

It would appear that most listings fall int he middle with a 'C' grade, 
although there are a significant number in all grades. 

When was Prosper's platform most popular? Is there a time when most of the 
listings were created?

```{r}
ggplot(aes(x=CreationMonth, group=1), data=data) + 
  geom_point(stat='count') + 
  geom_line(stat='count') + 
  scale_x_date(labels = date_format('%Y-%m'),
               date_breaks = '6 month',
               limits = c(as.Date('2005-6-1'), as.Date('2014-12-1'))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

There appears a sharp drop off in early to mid 2008 indicating an impact from
the recession. There is also a sharp drop off in early 2014, and it might be
useful to read up on the company and examine the possible cause. I don't see
a clear monthly pattern, but I'll check the distribution just in case. 


```{r}
ggplot(aes(x=NumericMonth), data=data) +
  geom_bar()

```

Next we'll check the frequency of different term lengths for the loans.

```{r}
ggplot(aes(x=Term), data=data) + geom_bar()

```

It looks like the vast majority of loans are 36 month (3 year) loans, with 
very few being a 12 month loan. It might be interesting later to look at the
relationship between loan term and the amount.


```{r}
ggplot(aes(x=BorrowerAPR), data=data) + geom_histogram(boundary=0)

```

```{r}
summary(data$BorrowerAPR)
```

The APR distribution appears to have a couple of distinct peaks with a few 
clear outliers at the high and low end. 

```{r}
ggplot(aes(x=ProsperRating), data=subset(data, !is.na(ProsperRating))) +
         geom_bar()
```

```{r}
ggplot(aes(x=ProsperScore), data=subset(data, !is.na(ProsperScore))) +
         geom_bar()
```

```{r}
summary(data$ListingCategory)
```

I want to look only at the most popular listing categories


```{r}
sub_data = subset(data, !is.na(ListingCategory) &
                     ListingCategory != 'Other' &
                     ListingCategory != 'Not Available')

categories = names(sort(table(sub_data$ListingCategory), decreasing = T)[1:5])

ggplot(aes(x=ListingCategory), 
       data=subset(data, ListingCategory %in% categories)) + 
  geom_bar()
```

Clearly Debt Consolidation is by far the most popular use for Prosper. 

```{r}
summary(data$StatedMonthlyIncome)
```

I would like to be the guy who makes 175K a month


```{r}
ggplot(aes(x=StatedMonthlyIncome), data=data) + 
  geom_histogram(boundary=0)
```

Hmm, I going to go ahead an ignore the top 1% of values because that is 
cleary skewing the distribution. 

```{r}
ggplot(aes(x=StatedMonthlyIncome), data=data) + 
  geom_histogram(boundary=0) + 
  xlim(0, quantile(data$StatedMonthlyIncome, .99))
```

It looks like either a log normal or exponential distribution. I'll take a 
log transformation to try and get a better idea of the data

```{r}
ggplot(aes(x=StatedMonthlyIncome), data=data) + 
  geom_histogram(binwidth = .1) + 
  xlim(0, quantile(data$StatedMonthlyIncome, .99)) + 
  scale_x_log10()
```

Now we're looking roughly normal with a few clear outliers and a low standard
deviation

I want to look into the high and low outliers that to see if there is 
anything weird going on. 

```{r}
head(data[order(data$StatedMonthlyIncome),])
```

A large number of people list their monthly income as 0, which screws with the
log distribution. After examining the orignal dataframe, it's difficult to tell 
whether this data is a mistake or not, so we'll just take it at face value. 

```{r}
tail(data[order(data$StatedMonthlyIncome),])
```




```{r}
ggplot(aes(x=LoanOriginalAmount), data=data) + 
  geom_histogram(binwidth = 3000, boundary=0)
```

Not a clear distribution for loan amount, but it appears as though it could be
normal. I'll look at the distribution of the log value just to see how it looks. 

```{r}
ggplot(aes(x=LoanOriginalAmount), data=data) + 
  geom_histogram(binwidth = .1) + 
  scale_x_log10()
```

Original loan amount doesn't really fit very neatly into a distribution. 


CreditScore:

```{r}
ggplot(aes(x=CreditScore), data=data) + geom_histogram()
```

Distribution appears roughly normal. It seems safe to assume that when the 
credit score is 0, that's a case where the data is simply incomplete. 

# Univariate Analysis

### What is the structure of your dataset?

There are 113,937 observations with 81 different features, however I chose to
for this analysis I made some basic transformations and chose to focus on 12 
of them. 

* CreditGrade
* ListingCreationDate 
* CreationMonth
* Term
* BorrowerAPR
* ProsperRating
* ProsperScore
* ListingCategory
* CreditScore
* StatedMonthlyIncome
* LoanOriginalAmount
* Investors

I chose to treat the variables Term, ProsperRating, and ProsperScore as ordered 
factors. The ordering is obvious as they are all whole numbers from 1 to the 
max level for the factor. ListingCategory is an unordered factor with 21 levels. 

Other observations: 

* Most loans have a small number of investors with 0 being most common. 
* Credit grades are pretty evenly distributed, although B,C, and D are more 
typical. 
* There was a drop off in listings with the recession and in general the service
has become more popular over time. 
* There is no pattern for creation month.
* The vast majority of loans have a 36 month term. 
* BorrowerAPR is normally distributed. 
* By far the most common listing category is "Debt Consolidation"

### What is/are the main feature(s) of interest in your dataset?

The main features of interest are BorrowerAPR and factors that intuitively 
correlate with risk, with the core one being CreditScore. 

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

I think that ProsperRating, ProsperScore, and CreditGrade are all important
factors in determining BorrowerAPR. I cannot say with certainty which will be 
most influential. 

### Did you create any new variables from existing variables in the dataset?

I averaged the high credit score and the low credit score in order to create a 
single numeric credit score. I also found the month for each listing creation 
date to look for any patterns. 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

A number of the features were clearly exponential distributions. When analyzing 
relationships, it is generally more informative when a variable is normally
distributed. For these variables I looked at the log of the distribution. Eg:

* Investors
* StatedMonthlyIncome (Also want to exclude very low values)
* LoanOriginalAmount

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}
sample = data[sample(nrow(data), 1000), ]
pairs(sample)
```

This is pretty noisy so I'll look into relationships with individual plots. 

First I want to look more closely at the impact of CreditGrade. For these plots
I'll exclude 'NA' and 'NC' grades, considering them incomplete data. 

```{r}
credit_grade_idx = (data$CreditGrade != 'NC') & (!is.na(data$CreditGrade))
ggplot(aes(x=CreditGrade, y=BorrowerAPR), data=data[credit_grade_idx, ]) + 
  geom_boxplot()
```

As expected, loans with better credit grades resulted in a lower rate for the 
borrower. However, for all cases there are borrowers that represent significant
outliers. I will look into these outlying observations later in the analysis. 


I also want to see if there is a relationship between the Term of the loan 
and the rate borrowers pay. Intuitively you might expect borrowers with longer
loan periods to pay a lower rate. 

```{r}
ggplot(aes(x=Term, y=BorrowerAPR), data=data) + 
  geom_boxplot()
```

It's interesting we see almost no difference in APR between different term 
periods. However, it is worth noting that the 60 month loans have a much lower
spread than the shorter loans. For concrete numbers here are the standard 
deviations: 

```{r}
apr_idx = !is.na(data$BorrowerAPR)
by(data[apr_idx,]$BorrowerAPR, data[apr_idx,]$Term, sd)
```

And interquartile ranges:

```{r}
by(data[apr_idx,]$BorrowerAPR, data[apr_idx,]$Term, IQR)
```

Prosper Rating: 

```{r}
rating_idx = !is.na(data$ProsperRating)
ggplot(aes(x=ProsperRating, y=BorrowerAPR), data=data[rating_idx,]) + 
  geom_boxplot()
```

As you might expect, higher ratings pay a lower APR and the differences appear
pretty significant. What is surprising is the huge number of outliers, 
especially for the lowest rating. 

ProsperScore:

```{r}
score_idx = !is.na(data$ProsperScore)
ggplot(aes(x=ProsperScore, y=BorrowerAPR), data=data[score_idx, ]) + 
  geom_boxplot()
```

The general pattern is consistent with what you might expect, but there is quite 
a bit more variability than with ProsperRating. Clearly many investors have 
different feelings about the level of risk assigned to loans than Prosper. 

ListingCategory: 

```{r}
listing_idx = (!is.na(data$ListingCategory) & 
                 (data$ListingCategory != 'Not Available'))
ggplot(aes(x=ListingCategory, y=BorrowerAPR), data=data[listing_idx, ]) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0))
```

In general, the listing category doesn't seem to have a big impact on the apr
the borrower pays. What interesting however are the two apparrent exceptions
with personal loans paying a lower apr and cosmetic procedures paying a higher
apr. Perhaps this says something about the biases of the lenders. 

StatedMonthlyIncome:

Next we'll look at a continuous variable. 

```{r}
income_idx = !is.na(data$StatedMonthlyIncome)
ggplot(aes(x=StatedMonthlyIncome, y=BorrowerAPR), 
       data=data[income_idx, ]) + 
  geom_point()
```

Recall from the histogram of income that a few very high values skew the 
distribution. I'll remove those to get a better idea of the relationship. 

```{r}
ggplot(aes(x=StatedMonthlyIncome, y=BorrowerAPR), 
       data=data[income_idx, ]) + 
  geom_point() + 
  xlim(0, quantile(data$StatedMonthlyIncome, .99))
```

A lot of incomes and APRs clearly are set to neat values. I'll jitter the points
to help get a clearer picture as well as increase the transparency. 

```{r}
income_log_idx = data$StatedMonthlyIncome > 1
ggplot(aes(x=StatedMonthlyIncome, y=BorrowerAPR), 
       data=data[income_idx & income_log_idx, ]) + 
  geom_point(alpha=.01, position = 'jitter') + 
  xlim(0, quantile(data$StatedMonthlyIncome, .99)) + 
  geom_smooth()
```

There doesn't appear to be any relationship at all between income and APR. If
anything there is a slight negative relationship. This is surprising to me, 
but I suspect those with larger incomes also ask for larger loan amounts. I 
would also like to try a simple log transformation, as earlier that helped 
the income distribution appear more normal. 

```{r}
ggplot(aes(x=StatedMonthlyIncome, y=BorrowerAPR), 
       data=data[income_idx, ]) + 
  geom_point(alpha=.01, position = 'jitter') + 
  #xlim(1, quantile(data$StatedMonthlyIncome, .99)) + 
  scale_x_log10(limits=c(quantile(data$StatedMonthlyIncome, .05), 
                         quantile(data$StatedMonthlyIncome, .99))) + 
  geom_smooth()
```

LoanOriginalAmount:

```{r}
amount_idx = !is.na(data$LoanOriginalAmount)
ggplot(aes(x=LoanOriginalAmount, y=BorrowerAPR), 
       data=data[amount_idx, ]) + 
  geom_point()
```

Again high density clusters of points and distinct verticals that are common. 
I'll fade and jitter the points. 

```{r}
ggplot(aes(x=LoanOriginalAmount, y=BorrowerAPR), 
       data=data[amount_idx, ]) + 
  #geom_point(alpha=.01, position = 'jitter_x')
  geom_jitter(alpha=.01, width=100) + 
  geom_smooth()
```

There is a large amount of variability, but there does appear to be a slight 
negative relationship between loan amount and borrower APR. I suspect this 
is a result of wealthier borrowers taking out larger loans. We'll check that
assumption quickly. 

```{r}
ggplot(aes(x=LoanOriginalAmount, y=StatedMonthlyIncome), 
       data=data) + 
  geom_point(position='jitter', alpha=.01) + 
  ylim(0, quantile(data$StatedMonthlyIncome, .99)) + 
  geom_smooth()
```

It's not perfect, but there's clearly a positionve linear relationship between
income and loan amount which supports my theory. 


Investors:

```{r}
investor_idx = !is.na(data$Investors)
ggplot(aes(x=Investors, y=BorrowerAPR), data=data[investor_idx, ]) + 
  geom_point()
```

In the univariate analysis it was clear that Investors follows an exponential
relationship so I'll take a log transformation. I also want to exclude 
cases with 0 investors. 

```{r}
ggplot(aes(x=Investors, y=BorrowerAPR), data=data[investor_idx, ]) + 
  geom_point(alpha=.05, position='jitter') + 
  scale_x_log10(limits=c(1, max(data$Investors)))
```

There isn't really a clear picture about the relationship between the number of
investors and the APR. 

CreditScore:
(Assuming anything under 300 is a mistake)

```{r}
credit_score_idx = !is.na(data$CreditScore) & (data$CreditScore >= 300)
ggplot(aes(x=CreditScore, y=BorrowerAPR), data=data[credit_score_idx, ]) + 
  geom_point(position = 'jitter', alpha=.01) + 
  geom_smooth(method='lm')
```

As expected, there appears to be a negative linear relationship between credit
score and APR. 


# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Unsurprisingly, the different grading systems (CreditGrade, ProsperRating, 
ProsperScore) were all strong predictors of BorrowerAPR, with ProsperScore
being the least reliable. 

The length of the loan (Term) does not appear to impact the BorrowerAPR. 

ListingCategory seems not to matter much with the notable exceptions of 
'Cosmetic Procedure' and 'Personal Loan'. 

After removing outliers, there appears to be a weak negative linear relationship
between StatedMonthlyIncome and BorrowerAPR. 

There is also a weak negative linear relationship between LoanOriginalAmount 
and BorrowerAPR. 

Also there is a clear negative lineral relationship between CreditScore and 
BorrowerAPR. 

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

There is a strong positive relationship between StatedMonthlyIncome and 
LoanOriginalAmount so I suspect having both in the model would be a solid 
predictor of BorrowerAPR. 

### What was the strongest relationship you found?

The relationship between CreditScore and BorrowerAPR. Intuitively, this makes 
a lot of sense. 

# Multivariate Plots Section

I want to explore the obvious outliers in my BorrowerAPR vs CreditGrade boxplots.
Why are some borrowers with good credit grades receiving high APRs and vice 
versa. 

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x=CreditScore, y=BorrowerAPR), 
       data=data[credit_score_idx & credit_grade_idx,]) + 
  geom_point(position='jitter', alpha=.1, aes(color=CreditGrade)) + 
  theme_gray() + 
  scale_color_brewer(type='seq')
```

Including CreditScore doesn't help us explain the CreditGrade APR outliers. It 
is clear from the clean color bands in the plot that CreditScore is almost 
entirely determinite of CreditGrade with the bands becoming a little fuzzier for
the lower credit scores. 

```{r}
sub_data = data[income_idx & income_log_idx & credit_grade_idx,]
ggplot(aes(x=StatedMonthlyIncome, y=BorrowerAPR), data=sub_data) + 
  geom_point(position='jitter', alpha=.3, aes(color=CreditGrade)) + 
  xlim(0, quantile(sub_data$StatedMonthlyIncome, .99)) + 
  scale_color_brewer(type='seq')

```

There actually doesn't appear to be a strong relationship between income and 
borrower APR at all. It might be more interesting to look at loan amount 
relative to income. 

```{r}
sub_data = data[income_idx & income_log_idx & credit_grade_idx,]
ggplot(aes(x=LoanOriginalAmount/StatedMonthlyIncome, y=BorrowerAPR), 
       data=sub_data) + 
  geom_point(position='jitter', alpha=.3, aes(color=CreditGrade)) + 
  xlim(0, 
       quantile(sub_data$LoanOriginalAmount/sub_data$StatedMonthlyIncome, .99))+
  scale_color_brewer(type='seq')
```

wh, so this is a little more informative. There appears to be a positive linear
relationship between this ratio and and APR. Let's limit ourselves ot just 
high credit grades. 

```{r}
idx = data$CreditGrade %in% c('AA', 'A')
sub_data = data[income_idx & income_log_idx & credit_grade_idx & idx,]
ggplot(aes(x=LoanOriginalAmount/StatedMonthlyIncome, y=BorrowerAPR), 
       data=sub_data) + 
  geom_point(position='jitter', alpha=.3, aes(color=CreditGrade)) + 
  xlim(0, 
       quantile(sub_data$LoanOriginalAmount/sub_data$StatedMonthlyIncome, .99)) +
  geom_smooth(method='lm')

```

It looks like the ratio of the loan to income helps to explain at least part of 
the reason borrowers with a high credit grade might be given a higher APR. This
is especially true for high credit grade borrowers. 

Let's also see if time when the loan is given has an impact on APR. 

```{r}
ggplot(aes(x=CreationMonth, y=BorrowerAPR), 
       data=data[credit_grade_idx,]) + 
  geom_point(position='jitter', alpha=.1, aes(color=CreditGrade))  
```

There's definitely a weird cap on higher APR rates for low credit grade
borrowers during the midst of the recession. 

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

It appears that a borrower's credit grade is almost entirely a function of their
averaged credit score. However, this doesn't entirely explain some of the 
weird outliers. One possible explantion appeared to be the loan to income 
ratio. 

### Were there any interesting or surprising interactions between features?

It was surprising how perfectly credit grade maps to credit score. I would have
thought Prosper used a more sophisticated mdoel. 

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

I used a basic linear regression to model the relationship between APR and
the loan amount to income ratio for high credit grade borrowers. 

------

# Final Plots and Summary


### Plot One
```{r echo=FALSE, Plot_One}
listing_idx = (!is.na(data$ListingCategory) & 
                 (data$ListingCategory != 'Not Available'))
ggplot(aes(x=ListingCategory, y=BorrowerAPR), data=data[listing_idx, ]) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0)) +
  ggtitle('Borrower APR by Listing Category')
```

### Description One

Although the listing category may not be the most informative feature for most 
loans, I though it was interesting because of the outliers. It's clear that the
lending market is not kind to those taking out loans for "cosmetic procedures", 
but there is often times a discount for something as generic as "personal use". 

### Plot Two
```{r echo=FALSE, Plot_Two}
credit_score_idx = !is.na(data$CreditScore) & (data$CreditScore >= 300)
ggplot(aes(x=CreditScore, y=BorrowerAPR), data=data[credit_score_idx, ]) + 
  geom_point(position = 'jitter', alpha=.01) + 
  geom_smooth(method='lm') + 
  ggtitle('Borrower APR vs Credit Score')

```

### Description Two

Unsurprisingly one of the best predictors of a borrower's APR is their credit 
score. It just demonstrates that even on more modern lending platforms, it's 
still important to carefully manage your credit. 


### Plot Three
```{r echo=FALSE, Plot_Three}
idx = data$CreditGrade %in% c('AA', 'A')
sub_data = data[income_idx & income_log_idx & credit_grade_idx & idx,]
ggplot(aes(x=LoanOriginalAmount/StatedMonthlyIncome, y=BorrowerAPR), 
       data=sub_data) + 
  geom_point(position='jitter', alpha=.3, aes(color=CreditGrade)) + 
  xlim(0, 
       quantile(sub_data$LoanOriginalAmount/sub_data$StatedMonthlyIncome, .99)) +
  geom_smooth(method='lm') + 
  xlab('Loan Amount to Income') +
  ggtitle('Borrower APR for top Credit Grades by Loan Amount to Income Ratio')
```

### Description Three

This plot shows the impact of a borrower's loan amount to income ratio on APR.
It demonstrates that people who borrow a lot realtive to their monthly income
can expect to pay a higher APR. 

------

# Reflection

The Prosper dataset contains over 100,000 loans with 81 different features for
each loan. In my project, I chose to focus on only a small subset of those 
features. First I plotted the distribution of each feature individually in order
to get a feel for my data and look for potential outliers or oddities. 

Next I looked into what factors determined the APR for an individual borrower. 
Unsurprisingly, I found the Prosper rating systems such as Credit Grade, Prosper
Rating, and Prosper Score were all strong indicators of what APR a borrower 
might pay. Additionally, credit score was strongly correlated with APR. 
However, there were numerous outliers and I wanted to examine how these
occurred. I found that borrowers with high credit grades who still borrowed a
large amount relative to their monthly incomes were likely to pay a higher APR 
than you might otherwise expect. 

There are obvious limitation with the conclusions I draw. First, I only looked 
at a very small subset of the total number of features. There could be much 
stronger correlations and patterns that I'm not seeing. Second, my final plots
only looked at Credit Grade, which only exists up till 2009, so I'm looking
at only a subset of my data. Given that there are so many variables that can 
affect APR, in the future I might look into outside economic indicators such
as Fed interest rates to see if those cause some of the noise in my data.